---
title: "Geo quality of GSA data"
author: "Lars Vilhuber"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    keep_md: yes
editor_options:
  chunk_output_type: console
---
We are only going to be looking at *civilian agencies*.

```{r setup, echo=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(knitr)
library(ggmap)
library(kableExtra)
opts_chunk$set(warning = FALSE)
opts_chunk$set(message=FALSE)
#opts_chunk$set(results='asis')
```

# Source
Data can be obtained from website at  https://www.gsa.gov/policy-regulations/policy/real-property-policy/asset-management/federal-real-property-profile-frpp/federal-real-property-public-data-set

## 2016 version
```{r, echo=FALSE}
loc.2016 <- "https://inventory.data.gov/dataset/70667db2-37b0-4a16-adda-beb476e3bf18/resource/e394b6f0-1888-4ffe-9ab1-bab83d2a95d3/download/fy-2016-civilian-agency-public-frpp-dataset.csv"
```
- [Location on Data.gov](https://catalog.data.gov/dataset/fy-2016-federal-real-property-profile-data-for-civilian-agencies/resource/adafcd00-cab8-424c-9ee2-6e45a97ba160)
- [Location elsewhere](https://d2d.gsa.gov/dataset/public-frpp-dataset-civilian-agencies)
- [Direct URL](`r loc.2016`)

## 2017 version
```{r,echo=FALSE}
loc.2017 <- "https://inventory.data.gov/dataset/509d3f36-464e-4e4c-a26b-8582605c254e/resource/37120970-e7cf-47e3-ba25-3f5a6b0f4c2f/download/fy-2017-civilian-agency-public-frpp-data.csv"
```
- [Location on Data.gov](https://catalog.data.gov/dataset/federal-real-property-profile-data-for-civilian-agencies/resource/cff1291c-547d-46fd-8827-47c57f2d8239)
- [Direct URL](`r loc.2017`)

# Characteristics:

-  Available Time Period:     October 1, 2015 to September 30, 2016
- Last Update:     March 16, 2018
-  Frequency of Updates:     Annual

```{r readin,error=FALSE}
fy_2016 <- read_csv("data/fy-2016-civilian-agency-public-frpp-dataset-12142017.csv",progress = FALSE)
# restrict to united states
fy_2016 <- subset(fy_2016,tolower(Country)=="united states")
```

The 2016 file has `r nrow(fy_2016)` observations in the  United States. The remainder are government installations abroad.

## Missing Geocodes
The file comes with latitude/ longitude, but for some reason, a significant portion are empty. Most of the records are for "owned" property:
```{r find_missing_geo}
fy_2016$missing_geo <- is.na(fy_2016$Latitude)
table(fy_2016$`Legal Interest Indicator`,fy_2016$missing_geo) %>% kable() %>%
	kable_styling("striped") %>%
	group_rows("Legal Interest",1,5) %>%
  add_header_above(c(" " = 1, "Missing Lat/Long" = 2))
```
It is not that the address is missing, although that contributes:
```{r find_missing_adr}
fy_2016$missing_adr <- is.na(fy_2016$`Street Address`)
table(fy_2016$missing_adr,fy_2016$missing_geo) %>% kable() %>%
	kable_styling("striped") %>%
	group_rows("Missing street address",1,2) %>%
  add_header_above(c(" " = 1, "Missing Lat/Long" = 2))
```
nor that the address cannot be geo-coded, when lat/long is missing but street address is available:
```{r setup_geocoding}
fy_2016$complete_address <- paste(fy_2016$`Street Address`,fy_2016$City,fy_2016$State,fy_2016$`ZIP Code`,sep=", ")
# we only run a sample of 100 records. The DataScienceToolkit uses OpenStreetMap for geocoding.
fy_2016_sample <- fy_2016 %>% 
	filter(missing_adr == FALSE, missing_geo == TRUE) %>%
	sample_n(100)
```
```{r geocode,results='hide'}
test_geo <- geocode(fy_2016_sample$complete_address,output=c("more"),source=c("dsk"))
```
```{r geocoding_results}
table(test_geo$type,test_geo$loctype) %>% kable() %>% 
	kable_styling("striped") %>%
	group_rows("Address precision",1,3) %>%
	add_header_above(c(" " = 1, "Geocoding precision" = 2))
```

# Coverage

The internal version of the FRPP also has the number of employees at the location. The public-use version only has square-footage. As a rough estimate of the fraction of population-weighted availability of geocoded data, we tabulate by square-footage (we neglect the distinction by "gross" and "rentable" square footage):
```{r sum_sqft}
fy_2016 %>% group_by(missing_geo) %>% summarize(`Square Feet`=sum(`Square Feet (Buildings)`,na.rm = TRUE)) %>%
	mutate("Percent" = `Square Feet`*100/sum(`Square Feet`)) %>%
	kable() %>%
	kable_styling("striped") 
	
```

